---
title: "Tidy Tuesday: 27 March Lab Meeting"
author: "AS Huffmyer"
date: "2023" 
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

Loading data for Tidy Tuesday lab meeting practice. 

# *Setup*  

Set knit options. 
```{r}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

Load packages. 

```{r}
library(tidyverse)
library(RColorBrewer)
```

Load data frame for FeederWatch

```{r}
# Read in the data manually

languages <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-03-21/languages.csv')

```

# *Examine dataset*  

Look at languages dataset. 
```{r}
str(languages)
```

# *Questions that I want to answer*  

1. Which languages have the most users? 
Plot: Bar plot with total users on the Y, language on the X, ordered by total users 

2. Does user count relate to the age of the language? 
Plot: Scatter plot with correlation with year created on the X and users on the Y 

3. Which languages are most highly used in GitHub repos? 
Plot: Bar plot with GitHub users on the Y, language on the X, ordered by GitHub users

4. Does having a language be open source relate to having more users? 
Plot: Scatter plot with language on X, users on Y, color by open source or not 

# 1. Which languages have the most users? 

First view a histogram of the number of users. 
```{r}
hist(languages$number_of_users)
```

We need to remove the lower user counts to be able to see more information about the higher user counts. Try removing any language with less than 100,000 users. 
```{r}
users<-languages%>%
  select(number_of_users)%>%
  filter(number_of_users>100000)

hist(users$number_of_users)
```

Plot the number of users for each language for languages with >100,000 users. 
```{r}
languages%>%
  filter(number_of_users>100000)%>%
  
  ggplot(aes(x=reorder(title, -number_of_users), y=number_of_users, fill=factor(ifelse(title=="R","Highlighted","Normal"))))+
  geom_bar(stat = "identity", color="black")+
  scale_fill_manual(name = "area", values=c("red","white")) +
  xlab("Language")+
  ylab("Number of Users")+
  scale_y_continuous(labels = scales::comma)+
  theme_classic()+
  theme(
    axis.text.x = element_text(angle=45, hjust=1),
    legend.position="none"
  )
```

The most used languages are SQL, Java, HTML, and C++. 

# 2. Does user count relate to the age of the language? 

Generate a scatter plot that shows the correlation between the year the language was created and the number of users. 

```{r}
languages%>%
  #filter(number_of_users>100000)%>%
  
  ggplot(aes(x=appeared, y=number_of_users))+
  geom_point(color="gray")+
  stat_smooth(color="black")+
  xlab("Year Appeared")+
  ylab("Number of Users")+
  scale_y_continuous(labels = scales::comma)+
  theme_classic()
```

This plot isn't that helpful. Instead plot relationship between year and number of users for languages with >100,000 users.  

```{r}
languages%>%
  filter(number_of_users>100000)%>%
  
  ggplot(aes(x=appeared, y=number_of_users, color=factor(ifelse(title=="R","Highlighted","Normal"))))+
  geom_point()+
  stat_smooth(method="lm", color="black")+
  geom_text(aes(label=ifelse(number_of_users>4000000,as.character(title),'')),hjust=-0.2,vjust=0, color="black")+
  geom_text(aes(label=ifelse(title=="R",as.character(title),'')),hjust=-0.2,vjust=0, color="red")+
  scale_color_manual(name = "area", values=c("red","gray")) +
  geom_text(aes(label=ifelse(appeared<1965,as.character(title),'')),hjust=-0.2,vjust=0, color="black")+
  geom_text(aes(label=ifelse(appeared>2010,as.character(title),'')),hjust=-0.2,vjust=0, color="black")+
  xlab("Year Appeared")+
  ylab("Number of Users")+
  scale_y_continuous(labels = scales::comma)+
  theme_classic()+
  theme(
    legend.position="none"
  )
```

There appears to be a weak positive relationship between age of lanugage and the number of users (newer programs tend to have fewer users).  

# 3. Which languages are most highly used in GitHub repos? 

First view a histogram of the number of GitHub repos 
```{r}
hist(languages$github_language_repos)
```

We need to remove the lower user counts to be able to see more information about the higher repo counts. Try removing any language with less than 100,000 repos. 
```{r}
git<-languages%>%
  select(github_language_repos)%>%
  filter(github_language_repos>100000)

hist(git$github_language_repos)
```

Plot the number of GitHub repos for each language for languages with >100,000 repos.  
```{r}
languages%>%
  filter(github_language_repos>100000)%>%
  
  ggplot(aes(x=reorder(title, -github_language_repos), y=github_language_repos, fill=factor(ifelse(title=="R","Highlighted","Normal"))))+
  geom_bar(stat = "identity", color="black")+
  scale_fill_manual(name = "area", values=c("red","white")) +
  xlab("Language")+
  ylab("Number of GitHub Repos")+
  scale_y_continuous(labels = scales::comma)+
  theme_classic()+
  theme(
    axis.text.x = element_text(angle=45, hjust=1),
    legend.position="none"
  )
```

The highest languages on GitHub are Java, HTML, and Python. 

# 4. Do open source languages have more users?

Plot the number of users for each language and color the correlations by those with open source programmings. 

Remove any languages that have NA in the `is_open_source` column that we don't have information for. 
```{r}
languages%>%
  filter(!is.na(is_open_source))%>%
  
  ggplot(aes(x=is_open_source, y=number_of_users, color=factor(ifelse(title=="R","Highlighted","Normal"))))+
  geom_point()+
  geom_text(aes(label=ifelse(number_of_users>500000,as.character(title),'')),hjust=-0.2,vjust=0, color="black")+
  geom_text(aes(label=ifelse(title=="R",as.character(title),'')),hjust=-0.2,vjust=0, color="red")+
  scale_color_manual(name = "area", values=c("red","gray")) +
  geom_text(aes(label=ifelse(is_open_source==FALSE & number_of_users>10000,as.character(title),'')),hjust=-0.2,vjust=0, color="black")+
  xlab("Is the language open source?")+
  ylab("Number of Users")+
  scale_y_continuous(labels = scales::comma)+
  theme_classic()+
  theme(
    legend.position="none"
  )
```

There are not many languages that we have information for from this data set and there are only a few that are not open source. Of those that we have information for, it appears that open source languages have more users. 